---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dev = "svg",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# echos

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Licence](https://img.shields.io/badge/licence-GPL--3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0.en.html)
[![Codecov test coverage](https://codecov.io/gh/ahaeusser/echos/branch/master/graph/badge.svg)](https://app.codecov.io/gh/ahaeusser/echos?branch=master)
[![R-CMD-check](https://github.com/ahaeusser/echos/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ahaeusser/echos/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The package provides a tidy interface for modeling and forecasting univariate time series using Echo State Networks (ESNs). The model works within the `fable` framework provided by the `fabletools` package, which provides the tools to evaluate, visualize, and combine models in a workflow consistent with the tidyverse.

***Disclaimer:*** The `echos` package is highly experimental and it is very likely that there will be (substantial) changes in the near future. These changes will probably affect the interface (e.g. arguments within `ESN()`) and the underlying modeling procedure itself.


## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("ahaeusser/echos")
```

## Getting started

```{r base}

library(echos)

# Forecast horizon
n_ahead <- 12 # forecast horizon
# Number of observations
n_obs <- length(AirPassengers)
# Number of observations for training
n_train <- n_obs - n_ahead

# Prepare train and test data
xtrain <- AirPassengers[(1:n_train)]
xtest <- AirPassengers[((n_train+1):n_obs)]

# Train and forecast ESN model
xmodel <- train_esn(y = xtrain)
xfcst <- forecast_esn(xmodel, n_ahead = n_ahead)

# Plot result
plot(xfcst, test = xtest)

```
