---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dev = "svg",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# echos

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Licence](https://img.shields.io/badge/licence-GPL--3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0.en.html)
[![Codecov test coverage](https://codecov.io/gh/ahaeusser/echos/branch/master/graph/badge.svg)](https://app.codecov.io/gh/ahaeusser/echos?branch=master)
<!-- badges: end -->

The package provides a tidy interface for modeling and forecasting univariate time series using Echo State Networks (ESNs). The model works within the `fable` framework provided by the `fabletools` package, which provides the tools to evaluate, visualize, and combine models in a workflow consistent with the tidyverse.

***Disclaimer:*** The `echos` package is highly experimental and it is very likely that there will be (substantial) changes in the near future. These changes will probably affect the interface (e.g. arguments within `ESN()`) and the underlying modeling procedure itself.


## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("ahaeusser/echos")
```

## Getting started

```{r base}

library(echos)

# Forecast horizon
n_ahead <- 24 # forecast horizon
# Number of observations
n_obs <- length(AirPassengers)
# Number of observations for training
n_train <- n_obs - n_ahead

# Prepare train and test data
xtrain <- AirPassengers[(1:n_train)]
xtest <- AirPassengers[((n_train+1):n_obs)]

# Train and forecast ESN model
xmodel <- train_esn(y = xtrain)
xfcst <- forecast_esn(xmodel, n_ahead = 24)

# Plot result
plot(xtest, type = "l")
lines(xfcst$point, col = "steelblue", lwd = 2)

```



<!-- ## Usage -->

<!-- ### Load packages -->
<!-- ```{r packages, warning=FALSE, message=FALSE} -->
<!-- library(echos) -->
<!-- library(tidyverse) -->
<!-- library(tsibble) -->
<!-- library(fabletools) -->
<!-- library(fable) -->
<!-- ``` -->

<!-- ```{r abbreviations, echo=FALSE, warning=FALSE, message=FALSE, results='hide'} -->
<!-- Sys.setlocale("LC_TIME", "C") -->
<!-- ``` -->


<!-- ### Prepare data -->

<!-- The dataset `m4_monthly` is a monthly `tsibble` contains the time series with the ID `"M10960"` from the M4 Forecasting Competition. The first 300 observations are used for model training and last 24 observations are used for testing (hold-out data). -->



<!-- ### Model -->

<!-- The function `fabletools::model()` is used to automatically train an ESN to the time series data. The object `mdl` is a `mable` containing the trained ESN. -->

<!-- ```{r model} -->
<!-- # Automatic model training of ESN -->
<!-- mdl <- data_train %>% -->
<!--   model("ESN" = ESN(value)) -->

<!-- mdl -->
<!-- ``` -->

<!-- The function `report` is used to get a detailed summary of the trained ESN. From the output below, you get the following information about the trained model: -->

<!-- * Network size -->
<!--   * `Inputs`: The number of input variables -->
<!--   * `Reservoir`: The number of internal states (the reservoir is the hidden layer of an ESN) -->
<!--   * `Outputs`: The number of output variables (= response variables) -->
<!-- * Model inputs (from the output, you can see that two model inputs are used (one lag plus the intercept term)) -->
<!--   * `Constant`: Indicates whether an intercept term is used or not -->
<!--   * `Lags`: The lags of the output variable, which are used as model input -->
<!-- * Differences: First differences are calculated to achieve stationarity -->
<!-- * Scaling -->
<!--   * `Inputs`: The training data are scaled to the interval `(-1, 1)` -->
<!--   * `Random uniform`: The input weights matrix and the reservoir weight matrix are drawn from a random uniform distribution with interval `(-0.5, 0.5)` -->
<!-- * Hyperparameters: -->
<!--   * `alpha`: Leakage rate (smoothing parameter) -->
<!--   * `rho`: Spectral radius for scaling the reservoir weight matrix -->
<!--   * `lambda`: Regularization parameter for the ridge regression -->
<!--   * `density`: The density of the reservoir weight matrix -->
<!-- * Metrics: -->
<!--   * `df`: Effective Degrees of Freedom -->
<!--   * `aic`: Akaike Information Criterion -->
<!--   * `bic`: Bayesian Information Criterion -->
<!--   * `hq`: Hannan-Quinn Information Criterion -->


<!-- ```{r summary} -->
<!-- # Detailed report of ESN -->
<!-- mdl %>% -->
<!--   select(ESN) %>% -->
<!--   report() -->

<!-- ``` -->


<!-- ### Forecast -->

<!-- The function `fabletools::forecast()` is used to forecast the trained model 24-steps ahead. The object `fcst` is a `fable` containing the forecasts of the ESN. -->

<!-- ```{r forecast} -->
<!-- # Forecast models -->
<!-- fcst <- mdl %>% -->
<!--   forecast(h = 24) -->

<!-- fcst -->
<!-- ``` -->


<!-- ### Visualize -->

<!-- Plot the forecast along the actual values (training and test data). -->

<!-- ```{r plot} -->
<!-- # Visualize forecast and actual values -->
<!-- fcst %>% -->
<!--   autoplot( -->
<!--     bind_rows( -->
<!--       data_train, -->
<!--       data_test)) -->

<!-- ``` -->


<!-- ## Work in Progress -->

<!-- * Implement further functions -->
<!--   + `refit.ESN()` -->
<!--   + `generate.ESN()` -->
<!--   + `stream.ESN()` -->
<!--   + `reservoir.ESN()` -->
<!--   + `plot_reservoir()` -->
<!--   + ... -->
<!-- * Enhance `ESN()` to multivariate time series -->
