---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dev = "svg",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# echos

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Licence](https://img.shields.io/badge/licence-GPL--3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0.en.html)
<!-- badges: end -->

The package provides a tidy interface for modeling and forecasting univariate time series using Echo State Networks (ESNs). The model works within the `fable` framework provided by the `fabletools` package, which provides the tools to evaluate, visualize, and combine models in a workflow consistent with the tidyverse.

***Disclaimer:*** The `echos` package is highly experimental and it is very likely that there will be (substantial) changes in the near future. These changes will probably affect the interface (e.g. arguments within `ESN()`) and the underlying modeling procedure itself.


## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("ahaeusser/echos")
```
## Usage

### Load packages
```{r packages, warning=FALSE, message=FALSE}
library(echos)
library(tscv)
library(dplyr)
library(tsibble)
library(fabletools)
library(fable)
```

```{r abbreviations, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
Sys.setlocale("LC_TIME", "C")
```


### Prepare data

The dataset `elec_price` is a hourly `tsibble` with day-ahead electricity spot prices in [EUR/MWh] from the ENTSO-E Transparency Platform. The dataset contains time series data from 2019-01-01 to 2019-12-31 for 8 european bidding zones (BZN). In this example, we will only model and forecast the data for Germany (`DE`) and for one specific split (`176`). The training data ranges from 2019-06-25 00:00:00 to 2019-10-02 23:00:00 and we want to forecast the next day (24-steps ahead) from 2019-10-03 00:00:00 to 2019-10-03 23:00:00.

```{r data}

# Prepare dataset
data <- elec_price %>%
  tscv::clean_data()

# Setup for time series cross validation
n_init <- 2400   # size for training window
n_ahead <- 24    # size for testing window (forecast horizon)
mode <- "slide"  # fixed window approach
n_skip <- 23     # skip 23 observations
n_lag <- 0       # no lag

data <- data %>%
  tscv::split_data(
    n_init = n_init,
    n_ahead = n_ahead,
    mode = mode,
    n_skip = n_skip,
    n_lag = n_lag)

# Use only a single time series
data <- data %>%
  filter(BZN == "DE") %>%  # filter to bidding zone Germany
  filter(split == 176)     # filter to split 176

# Use only training data for modeling
data_train <- data %>%
  filter(sample == "train")

```

### Model

The function `fabletools::model()` is used to automatically train an ESN to the time series data. The object `mdl` is a `mable` containing the trained ESN.

```{r model}
# Automatic model training of ESN
mdl <- data_train %>%
  fabletools::model("ESN" = ESN(Value))

mdl
```

The function `report` is used to get a detailed summary of the trained ESN. From the output below, you get the following information about the trained model:

* Network size
  * `Inputs`: The number of input variables
  * `Reservoir`: The number of internal states (the reservoir is the hidden layer of an ESN)
  * `Outputs`: The number of output variables (= response variables)
* Model inputs (from the output, you can see that four model inputs are used (three lags plus the intercept term))
  * `Constant`: Indicates whether an intercept term is used or not
  * `Lags`: The lags of the output variable, which are used as model input
* Differences: First differences are calculated to achieve stationarity
* Scaling
  * `Inputs`: The training data are scaled to the interval `(-1, 1)`
  * `Random uniform`: The input weights matrix and the reservoir weight matrix are drawn from a random uniform distribution with interval `(-0.5, 0.5)`
* Hyperparameters:
  * `alpha`: Leakage rate (smoothing parameter)
  * `rho`: Spectral radius for scaling the reservoir weight matrix
  * `lambda`: Regularization parameter for the ridge regression
  * `density`: The density of the reservoir weight matrix
* Metrics:
  * `df`: Effective Degrees of Freedom
  * `aic`: Akaike Information Criterion
  * `bic`: Bayesian Information Criterion
  * `hq`: Hannan-Quinn Information Criterion


```{r summary}
# Detailed report of ESN
mdl %>%
  select(ESN) %>%
  report()

```


### Forecast

The function `fabletools::forecast()` is used to forecast the trained model 24-steps ahead. The object `fcst` is a `fable` containing the forecasts of the ESN.

```{r forecast}
# Forecast models
fcst <- mdl %>%
  fabletools::forecast(h = n_ahead)

fcst
```


### Visualize

Plot the forecast along the actual values.

```{r plot}
tscv::plot_forecast(
  fcst = fcst,         # forecasts as fable
  data = data,         # training and test data
  split = 176,         # only one split is visualized
  include = 72)        # plot only the last 72 observations of training data

```


## Work in Progress

* Implement further functions
  + `refit.ESN()`
  + `generate.ESN()`
  + `stream.ESN()`
  + `reservoir.ESN()`
  + `plot_reservoir()`
  + ...
* Enhance `ESN()` to multivariate time series
