---
title: "Introduction"
author: "Alexander Häußer"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height = 5,
  fig.width = 7
)
```

The package provides a tidy interface for modeling and forecasting univariate time series using Echo State Networks (ESNs). The model works within the `fable` framework provided by the `fabletools` package, which provides the tools to evaluate, visualize, and combine models in a workflow consistent with the tidyverse.

***Disclaimer:*** The `echos` package is highly experimental and it is very likely that there will be (substantial) changes in the near future. These changes will probably affect the interface (e.g. arguments within `ESN()`) and the underlying modeling procedure itself.


## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("ahaeusser/echos")
```
## Usage

### Load packages
```{r packages, warning=FALSE, message=FALSE}
library(echos)
library(tidyverse)
library(tsibble)
library(fabletools)
library(fable)
```

```{r abbreviations, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
Sys.setlocale("LC_TIME", "C")
```


### Prepare data

The dataset `m4_monthly` is a monthly `tsibble` contains the time series with the ID `"M10960"` from the M4 Forecasting Competition. The first 300 observations are used for model training and last 24 observations are used for testing (hold-out data).

```{r data}
# Split data into training and testing
data_train <- m4_monthly %>%
  slice_head(n = 300)

data_train


data_test <- m4_monthly %>%
  slice_tail(n = 24)

data_test

```

### Model

The function `fabletools::model()` is used to automatically train an ESN to the time series data. The object `mdl` is a `mable` containing the trained ESN.

```{r model}
# Automatic model training of ESN
mdl <- data_train %>%
  model("ESN" = ESN(value))

mdl
```

The function `report` is used to get a detailed summary of the trained ESN. From the output below, you get the following information about the trained model:

* Network size
  * `Inputs`: The number of input variables
  * `Reservoir`: The number of internal states (the reservoir is the hidden layer of an ESN)
  * `Outputs`: The number of output variables (= response variables)
* Model inputs (from the output, you can see that two model inputs are used (one lag plus the intercept term))
  * `Constant`: Indicates whether an intercept term is used or not
  * `Lags`: The lags of the output variable, which are used as model input
* Differences: First differences are calculated to achieve stationarity
* Scaling
  * `Inputs`: The training data are scaled to the interval `(-1, 1)`
  * `Random uniform`: The input weights matrix and the reservoir weight matrix are drawn from a random uniform distribution with interval `(-0.5, 0.5)`
* Hyperparameters:
  * `alpha`: Leakage rate (smoothing parameter)
  * `rho`: Spectral radius for scaling the reservoir weight matrix
  * `lambda`: Regularization parameter for the ridge regression
  * `density`: The density of the reservoir weight matrix
* Metrics:
  * `df`: Effective Degrees of Freedom
  * `aic`: Akaike Information Criterion
  * `bic`: Bayesian Information Criterion
  * `hq`: Hannan-Quinn Information Criterion


```{r summary}
# Detailed report of ESN
mdl %>%
  select(ESN) %>%
  report()

```


### Forecast

The function `fabletools::forecast()` is used to forecast the trained model 24-steps ahead. The object `fcst` is a `fable` containing the forecasts of the ESN.

```{r forecast}
# Forecast models
fcst <- mdl %>%
  forecast(h = 24)

fcst
```


### Visualize

Plot the forecast along the actual values (training and test data).

```{r plot}
# Visualize forecast and actual values
fcst %>%
  autoplot(
    bind_rows(
      data_train,
      data_test))

```
